---
title: "Visualising the Covid-19 Case Load for South Africa"
author: "Martin J Page"
date: "26/05/2020"
output: slidy_presentation

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Covid-19 in South Africa

- Covid-19 is a rapidly developing situation across the world  
- Citizens are interesting in the number of cases in their local areas
- This app allows users to view the total number of cases per province in South Africa between 24 March 2020 and 24 May 2020


```{r, warning=FALSE, error=FALSE, message=FALSE}
  library(leaflet)
  library(readr)
  library(dplyr)

    url <- "https://raw.githubusercontent.com/martinjpage/datasciencecoursera/master/RSACovid.csv"
    data <- read_csv(url)
    
    df <- data %>% group_by(Provinces)
    
      leaflet() %>% addTiles() %>% 
            addMarkers(lat = df$latit, lng = df$long,
                             label  = as.character(df$Provinces))

```



## Collecting the Data

- The data were programatically scraped from situation reports published on https://sacoronavirus.co.za/, a government website 
- The code for the webscraping is available at https://github.com/martinjpage/datasciencecoursera/blob/master/webscaping
- This table was joined with a table of the coordiates of the capital cities of each province
- The final data was read to an output file available at: https://github.com/martinjpage/datasciencecoursera/blob/master/RSACovid.csv
- Example of the final data:

```{r, warning=FALSE, error=FALSE, message=FALSE}
    library(readr)
    url <- "https://raw.githubusercontent.com/martinjpage/datasciencecoursera/master/RSACovid.csv"
    data <- read_csv(url)
    head(data)
```


## The App: How it Works
- The app is very easy to use
- Users select a date from an input calendar of available dates
- After clicking submit, the map will update to display the total case load per province for that day
- Documentation for using the app is provided in the sidebar
- The map is rendered using Leaflet
- Extract of plotting code from the server:

```{r, echo=TRUE, eval=FALSE}
output$plot1 <- renderLeaflet({
        df <- data %>% filter(Date == input$date)
        leaflet() %>% addTiles() %>% 
            addCircleMarkers(lat = df$latit, lng = df$long,
                             label  = as.character(df$Total), 
                             radius = sqrt(df$Total), 
                             color = RColorBrewer::brewer.pal(9, "Set1"), opacity = 0.8)
    })
```



## Check it out for yourself

- View the app at https://martinjpage.shinyapps.io/CovidRSA2/
- Example of the output plot with the date selected as 17 April 2020

```{r, message=FALSE, warning=FALSE}
    library(dplyr)
    library(readr)
    library(leaflet)

    url <- "https://raw.githubusercontent.com/martinjpage/datasciencecoursera/master/RSACovid.csv"
    data <- read_csv(url)
    
     df <- data %>% filter(Date == "2020-04-17")
        leaflet() %>% addTiles() %>% 
            addCircleMarkers(lat = df$latit, lng = df$long,
                             label  = as.character(df$Total), 
                             radius = sqrt(df$Total), 
                             color = RColorBrewer::brewer.pal(9, "Set1"), opacity = 0.8)
```


